Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System

### 2016-11-24 manuscript.tex
20161124

\begin{abstract}
This paper presents a novel adaptive dynamic sliding mode scheme for the deployment of the tethered satellite system. To overcome the limited inputs of the dynamic model, a new scalable dimensionless transformation and strictly bounded terms are designed, which guarantee the existence of the ratio between limited inputs and command signals. The adaptive saturated sliding mode control scheme is proposed to govern the tether deployment by introducing the adaptive rate into the system gain. The stability analysis indicates that this measure can eliminate the uncertainty caused by the input limitation and guarantee the asymptotically stability of the deployment dynamics. Finally, numerical simulations are presented to verify the effectiveness of the proposed control laws.
\end{abstract}
\begin{keyword}
Tethered satellite system;  Sliding mode control;  Input limitation; Adaptive control
\end{keyword}
\end{frontmatter}
\section{Introduction}
The deployment is the key technology of the space tethered satellite system (TSS)~\cite{wen2008advances,yousefian2015anti}, which supports various applications for space explorations and scientific researches, such as space debris removals, space elevators, satellite de-orbits and the tethered space robots~\cite{zhao2014thrust,kojima2015mission,zhong2013dynamics,zhong2013long,huang2014optimal,Huang2016Impact,Huang201637,Huang2016}. Many international space research institutes have launched orbital, suborbital and ground experiments, such as TSS-1R, SEDS, MAST and YES2, in an effort to verify the feasibility of the proposed theoretical and simulation schemes for the deployment of the TSS~\cite{williams2012review,robitaille2006interpreting,stone1998tss,williams2009yes2}. These experimental results intensively help researchers proceed to promote advanced mechanics, kinetic and control studies, to tackle the future challenges of space activities~\cite{yu2016nonlinear,wen2015space,aslanov2016swing,meng2016lyapunov,zhang2015line,Xu2016Hybrid}.

@@

for a vector $\bm x=\left(x_1,x_2,,\cdots,x_n\right)^T$. $\alpha$ and $\gamma$ are control law parameters which will be discussed in the control law design. Define the tracking error as $\Delta \bm q=\bm q-\bm q_d$, in which $\bm q_d$ is the desired trajectory with $\bm q_d(t)\in C^2$. However, the PID controller~(\ref{eq:pid controller}) can not be directly applied to the system~(\ref{eq:dynamic model 2}) with the input limitation, and in the subsequent section, we will utilize the saturated sliding mode technique to modify the controller~(\ref{eq:pid controller}) for the system~(\ref{eq:dynamic model 2}). For deriving the proposed control scheme, some properties should be given out in advance.

@@

\bm S_{q1} &= \mathbf{Sine}(\Delta \dot {\bm q}+\alpha\Delta\bm q),

@@

\bm S_{q2} &= \Delta \dot {\bm q}+\alpha \Delta \bm q,

@@

@@

In the aforementioned controller design, signum functions in inputs lead to chattering phenomena, which may cause adverse unmodeled dynamics to ruin the control system. Herein, we adopt the boundary layer technique to reduce chattering through replacing $\rm{sign}\it{(x)}$ by $\rm{sat}\it{(x/\nu)}$, where

@@

\begin{mythm}
Considering the controller in Theorem~\ref{thm:1}, we replace $\mathbf{sign}{(\bm{x})}$ in Eq.~(\ref{eq:slider control}) by $\mathbf{sat}{(\boldsymbol {x/\nu})}$, where $\mathbf{sat}{(\boldsymbol{x}/\boldsymbol{\nu})}=\left(\rm{sat}\it{(x_1/\nu_1)},\rm{sat}\it{(x_2/\nu_2)},\cdots,\rm{sat}\it{(x_n/\nu_n)}\right)^T$, to remove the chattering caused by the signum function, and the modified controller can guarantee the precisions of system states $\Delta\bm q$ that satisfy

@@

  \Vert\Delta\bm q\Vert\le\frac{\Vert\boldsymbol{arcsin}(\bm \nu)\Vert}{\alpha}.

@@

$\bm z$ is the position vector that describes the distances of system states in the boundary layer from the sliding mode surface, and $\dot{\bm z}$ is the velocity vector correspondingly. It is obvious that $\bm z$ is bounded and has $\Vert\bm z\Vert\le\Vert\bm\nu\Vert$, where $\bm\nu=(\nu_1,\nu_2,\nu_3)^T,\nu_1=\nu_2=\nu_3\ll 1$ is a constant vector denoting the boundary layer thicknesses for the three channels. Since all the inputs are bounded, it is evident to assume that $\dot{\bm z}$ is bounded with $\Vert\dot{\bm z}\Vert\le \bar z\le 2\Vert\bm\nu\Vert/\rm{d}\it{t}$.

@@

The selected Lyapunov function $V_3=\frac{1}{2}\bm{ S}_{q2}^T\bm{ S}_{q2}$ gives rise to

@@

&\le -\frac{\gamma \Vert \bm S_{q1}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\Vert\dot{\bm{z}}\Vert(\Vert\bm z\Vert+\Vert\mathbf{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma})\Vert)\\
&\le -\frac{\gamma \Vert \bm S_{q1}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\bar z(\Vert\bm{\nu}\Vert+\pi/2)\\

@@

Based on the Lyapunove stability theory, the above expression implies that $\bm S_{q2} = \Delta \dot {\bm q}+\alpha \Delta \bm q=0$ can attract the system states under the condition

@@

  \Vert \bm S_{q1}\Vert\ge\sqrt{\frac{\bar z(\Vert\bm{\nu}\Vert+\pi/2)}{\gamma}}
\end{align} Rephrasing that, the controller in Theorem~\ref{thm:1} with boundary layer technique can guarantee $\bm S_{q1}$ to converge to $\Vert\bm S_{q1}\Vert\le\sqrt{\frac{\bar z(\Vert\bm{\nu}\Vert+\pi/2)}{\gamma}}$, and once the states reach the boundary layer, since $\bm S_{q1}=\mathbf{Sine}(\Delta\dot{\bm q}+\alpha\Delta\bm q)$ we can combine

@@

  \mathbf{Sine}(\Delta\dot{\bm q}+\alpha\Delta\bm q)=\bm\nu

@@

with the Lyapunov function $V_4=\frac{1}{2}\Delta\bm q^T\Delta\bm q$ to obtain

@@

&=\Delta\bm q^T(-\alpha\Delta\bm q+\boldsymbol{arcsin}(\bm \nu))\\
&\le-\alpha\Vert\Delta\bm q\Vert^2+\Vert\Delta\bm q\Vert\Vert\boldsymbol{arcsin}(\bm \nu)\Vert\\
&\le-(\alpha\Vert\Delta\bm q\Vert-\Vert\boldsymbol{arcsin}(\bm \nu)\Vert)\Vert\Delta\bm q\Vert,

@@

which indicates that $\Vert\Delta\bm q\Vert=\frac{\Vert\boldsymbol{arcsin}(\bm \nu)\Vert}{\alpha}$ is an attractor, and hence, the ultimately system state precision of the controller with boundary layer technique is
  \Vert\Delta\bm q\Vert\le\frac{\Vert\boldsymbol{arcsin}(\bm \nu)\Vert}{\alpha}.
This completes the proof.

@@

@@

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\end{figure}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}

>>>>

\begin{abstract}
This paper presents a novel adaptive dynamic sliding mode scheme for the deployment of the tethered satellite system. To overcome the limited inputs of the dynamic model, a new scalable dimensionless transformation and strictly bounded terms are designed, which guarantee the existence of the ratio between limited inputs and command signals. The adaptive saturated sliding mode control scheme is proposed to govern the tether deployment by introducing the adaptive rate into the system gain. The stability analysis indicates that this measure can eliminate the uncertainty caused by the input limitation and guarantee the asymptotically stability of the deployment dynamics. Finally, numerical simulations are presented to verify the effectiveness of the proposed control laws.
\end{abstract}
\begin{keyword}
Tethered satellite system;  Sliding mode control;  Input limitation; Adaptive control
\end{keyword}
\end{frontmatter}
\section{Introduction}
The deployment is the key technology of the space tethered satellite system (TSS)~\cite{wen2008advances,yousefian2015anti}, which supports various applications for space explorations and scientific researches, such as space debris removals, space elevators, satellite de-orbits and the tethered space robots~\cite{zhao2014thrust,kojima2015mission,zhong2013dynamics,zhong2013long,Huang2016Impact,Huang2016,Huang2016416,Huang2016171,Huang20161985}. Many international space research institutes have launched orbital, suborbital and ground experiments, such as TSS-1R, SEDS, MAST and YES2, in an effort to verify the feasibility of the proposed theoretical and simulation schemes for the deployment of the TSS~\cite{williams2012review,robitaille2006interpreting,stone1998tss,williams2009yes2}. These experimental results intensively help researchers proceed to promote advanced mechanics, kinetic and control studies, to tackle the future challenges of space activities~\cite{yu2016nonlinear,wen2015space,aslanov2016swing,meng2016lyapunov,zhang2015line,Xu2016Hybrid}.

@@

for a vector $\bm x=\left(x_1,x_2,,\cdots,x_n\right)^T$. $\alpha$ and $\gamma$ are control law parameters which will be discussed in the control law design. Define the system states error as $\Delta \bm q=\bm q-\bm q_d$, in which $\bm q_d$ is the desired trajectory with $\bm q_d(t)\in C^2$. However, the PID controller~(\ref{eq:pid controller}) can not be directly applied to the system~(\ref{eq:dynamic model 2}) with the input limitation, and in the subsequent section, we will utilize the saturated sliding mode technique to modify the controller~(\ref{eq:pid controller}) for the system~(\ref{eq:dynamic model 2}). For deriving the proposed control scheme, some properties should be given out in advance.

@@

\bm S_{q1} &= \mathbf{Sine}(\Delta \dot {\bm q}+\alpha\Delta\bm q),\label{eq:sq1}

@@

\bm S_{q2} &= \Delta \dot {\bm q}+\alpha \Delta \bm q,\label{eq:sq2}

@@

  \end{cases}\label{eq:gamma}

@@

\textcolor{red}{In the aforementioned controller design, signum functions in inputs lead to chattering phenomena, which may cause adverse unmodeled dynamics to ruin the control system. Herein, we adopt the boundary layer technique to reduce chattering through replacing $\rm{sign}\it{(x)}$ by $\rm{sat}\it{(x/\nu)}$, where

@@

\begin{mythm}\label{thm:boundary}
Considering the controller in Theorem~\ref{thm:1}, we replace $\mathbf{sign}{(\bm{x})}$ in Eq.~(\ref{eq:slider control}) by $\mathbf{sat}{(\boldsymbol {x/\nu})}$, where $\mathbf{sat}{(\boldsymbol{x}/\boldsymbol{\nu})}=\left(\rm{sat}\it{(x_1/\nu_1)},\rm{sat}\it{(x_2/\nu_2)},\cdots,\rm{sat}\it{(x_n/\nu_n)}\right)^T$, to remove the chattering caused by the signum function, and the modified controller can guarantee that the system state error $\Delta\bm q$ is bounded on the boundary of the set $D_1^c$, which is the complement of $D_1$ defined as

@@

  D_1\triangleq\left\{\left(\Delta\bm q,\bar z,\gamma,\eta,\alpha\right)\Big\vert\Vert\Delta\bm q\Vert-\frac{\bar z}{\alpha\gamma\eta^2}\le 0\right\}.

@@

$\bm z$ is the position vector that describes the distances of system states in the boundary layer from the sliding mode surface, and $\dot{\bm z}$ is the velocity vector correspondingly. It is obvious that $\bm z$ is bounded and has $\Vert\bm z\Vert\le\Vert\bm\nu\Vert$, where $\bm\nu=(\nu_1,\nu_2,\nu_3)^T,\nu_1=\nu_2=\nu_3\ll 1$ is a constant vector denoting the boundary layer thicknesses for the three channels. Since all the inputs are bounded, it is evident to assume that $\dot{\bm z}$ is bounded with $\Vert\dot{\bm z}\Vert\le \bar z\le 2\Vert\bm\nu\Vert/\rm{d}\it{t}$. $\eta$ is defined as a time dependent positive ratio $\eta$ that depicts the relation $\Vert\bm S_{q1}\Vert=\eta\Vert\bm S_{q2}\Vert$, for all $\Vert\bm S_{q1}\Vert\neq 0$, $0<\eta<1$, when $\Vert\bm S_{q1}\Vert= 0$, $\eta$ especially values as $\eta = 1$.

@@

Choosing Lyapunov function $V_3=\frac{1}{2}\bm{ S}_{q2}^T\bm{ S}_{q2}$, its time derivative is given by

@@

&\le -\frac{\gamma \Vert \bm S_{q1}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\Vert\dot{\bm{z}}\Vert\Vert\bm{ S}_{q2}\Vert.\label{eq:v3}

@@

Reviewing Eq.~(\ref{eq:sq1}) and Eq.~(\ref{eq:sq2}), we get

@@

\bm S_{q1}=\mathbf{Sine}(\bm S_{q2})=\mathbf{Sine}(\Delta\dot{\bm q}+\alpha\Delta\bm q)
and since $\Vert\bm S_{q1}\Vert=\eta\Vert\bm S_{q2}\Vert$, Eq.~(\ref{eq:v3}) can be rewritten as

@@

  \dot V_3\le -\frac{\gamma\eta^2 \Vert \bm S_{q2}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\bar{z}\Vert\bm{ S}_{q2}\Vert

@@

Based on LaSalle theorem~\cite{lasalle1960some}, the above expression implies that $\bm S_{q2}$ can be attracted to the boundary of $D_2^c$, which is the complement of the set $D_2$ defined as
  D_2\triangleq\left\{(\bm S_{q2},\bar z,\gamma,\eta)\Big\vert \frac{\bar z}{\gamma\eta^2}-\Vert \bm S_{q2}\Vert\ge 0\right\}
Rephrasing that, the controller in Theorem~\ref{thm:1} with boundary layer technique can drive $\bm S_{q2}$ to the boundary of $D_2^c$. Considering the situation that the system state error maintains on the boundary of $D_2^c$, we can accordingly analyze the ultimate precision of system state error $\Delta\bm q$ by
&\Delta \dot {\bm q}+\alpha\Delta\bm q=\hat{\bm z}\\
&\Vert\hat{\bm z}\Vert=\frac{\bar z}{\gamma\eta^2}.
Choosing the Lyapunov function $V_4=\frac{1}{2}\Delta\bm q^T\Delta\bm q$, its time derivative is given by

@@

&=\Delta\bm q^T(-\alpha\Delta\bm q+\hat{\bm z})\\
&\le-\alpha\Vert\Delta\bm q\Vert^2+\Vert\Delta\bm q\Vert\Vert\hat{\bm z}\Vert\\
&\le-(\alpha\Vert\Delta\bm q\Vert-\Vert\hat{\bm z}\Vert)\Vert\Delta\bm q\Vert,

@@

Hence, by LaSalle theorem~\cite{lasalle1960some}, it follows that $\Delta\bm q$ is bounded on the boundary of the set $D_1^c$, which is the complement of $D_1$. This completes the proof.

@@

}
Particularly, in this work, the boundary layer is always attractive and is designed to be very thin. Moreover, by utilizing the proposed controller, the bounded inputs ensure that the dynamics in boundary layer are windless, which means that the change of $\bm z$ is not drastic and makes $\dot{\bm z}$ keep in a low level. We thus reasonably assume that $\bar z$ is a tiny positive scalar to ensure that $D_1$ is a small vicinity of origin. The proposed controller works to decrease the system state errors $\Delta\bm q$ and $\Delta\dot{\bm q}$, and this decreasing process tends to increasing $\eta$ contrarily. Hence, we can conclude that the system state error can maintain in a low level after reaching the boundary layer.

@@

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. \textcolor{red}{The control precisions in Figs~\ref{fig:7}-\ref{fig:9} are almost same to Figs~\ref{fig:3}-\ref{fig:6}, which means the loss of precision caused by the boundary layer technique is limited. Moreover, the simulation results show that the good performance of proposed controller and the suitable boundary layer thickness guarantee that the system state error converges to the small vicinity of origin, which is consistent with the analyses in Theorem~\ref{thm:boundary}.} Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\end{figure}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}
